A Bioconductor Workflow for Processing and Analysing Spatial Proteomics Data

### 2016-06-12 bioc-workflow.Rmd
some changes to QC section

Data quality is routinely examined through visualisation to verify that sub-cellular niches have been separated along the gradient. Based on De Duve's principle [@DeDuve:1981] proteins that co-localise exhibit similar quantitation profiles across the gradient fractions employed. One approach that has been widely used to visualise and inspect high throughput mass spectrometry-based proteomics data is principal components analysis (PCA). PCA is one of many dimensionality reduction methods, that allow one to effectively summarise multi-dimensional data in to 2 or 3 dimensions to enable visualisation. Very generally, the original continuous multi-dimensional data is transformed into a set of orthogonal components ordered according to the amount of variability that they describe. The `plot2D` method in `r Biocpkg("pRoloc")` allows one to plot the principal components (PCs) of a dataset against one another, by default the first two components are plotted on the x- and y-axis, respectively (the `dims` argument can be used to plot other PCs). If distinct clusters are observed, we assume that there is organellar separation present in the data. In the code chunk below we produce a PCA plot of the mouse stem cell dataset. One point on the plot represents one protein. We can indeed see several distinct protein clusters. Although, plotting the PCs does not give us a hard quantitative measure of separation, it is extremely useful summarising complex experimental information in one figure, to get an simplified overview of the data.

@@

**TODO** Mention that there are other dimensionality reduction
techniques. Advantages, disadvantages? See
[this](https://gist.github.com/mikelove/74bbf5c41010ae1dc94281cface90d32)
very interesting gist with intriguing tSNE examples.

@@

>>>>

Data quality is routinely examined through visualisation to verify
that sub-cellular niches have been separated along the gradient. Based
on De Duve's principle [@DeDuve:1981] proteins that co-localise
exhibit similar quantitation profiles across the gradient fractions
employed. One approach that has been widely used to visualise and
inspect high throughput mass spectrometry-based proteomics data is
principal components analysis (PCA). PCA is one of many dimensionality
reduction methods, that allow one to effectively summarise
multi-dimensional data in to 2 or 3 dimensions to enable
visualisation. Very generally, the original continuous
multi-dimensional data is transformed into a set of orthogonal
components ordered according to the amount of variability that they
describe. The `plot2D` method in `r Biocpkg("pRoloc")` allows one to
plot the principal components (PCs) of a dataset against one another,
by default the first two components are plotted on the x- and y-axis,
respectively (the `dims` argument can be used to plot other PCs). If
distinct clusters are observed, we assume that there is organellar
separation present in the data. In the code chunk below we produce a
PCA plot of the mouse stem cell dataset. One point on the plot
represents one protein. We can indeed see several distinct protein
clusters. Although, representing the multi-dimensional data along a
limited set of PCs does not give us a hard quantitative measure of
separation, it is extremely useful summarising complex experimental
information in one figure, to get an simplified overview of the data.

In the code chunk below, we specify `fcol = NULL`, which means not to
consider any feature variable to annotate the features with
colours. We will see later how to use this to annotate the PCA plot
with prior information about sub-cellular markers.

@@

We advise to visualise the without any annotation (i.e. with `fcol =
NULL`) in a first place, before proceeding with data annotation. The
identification of well resolved clusters in the data, constitutes an
unbiased assessment of the data structure, demonstrating the
successful separation of sub-cellular clusters.

There are other dimensionality reduction methods available in the
`plot2D` function, which can be parametrised with the `method`
argument.

**TODO** Mention that there are other dimensionality reduction
techniques. Advantages, disadvantages? See
[this](https://gist.github.com/mikelove/74bbf5c41010ae1dc94281cface90d32)
very interesting gist with intriguing tSNE examples. This might be
better to come after markers.

@@

A section about  `r Biocpkg("pRolocGUI")`.
